Disclosure of Product System Models in Life Cycle Assessment: Achieving Transparency and Privacy

### 2017-11-15 revision/reviewer-response.txt

Again, I have tried to take this advice to heart, but to some extent I have reduced the scope of the manuscript to exclude some of these concerns.  Again the section on "Stable semantic references" is meant to be a call to the community to consider precisely these problems and develop collective solutions.

+ The expected consequences for data providers, data users, database providers, and software developers should be mapped out in more detail. Where are the bottlenecks in setting up this process?

In the revised manuscript, these implications are no longer in scope.  The idea of structured disclosure of product system models is directed at modellers and report authors; it will ultimately bear on software makers but that is a matter to be addressed later.

Minor comments:

+ Please also refer to the allocation problem (co-production) and to the crediting problem (material or waste enter different product systems), and cite some relevant literature here.

Factors such as co-production strategy (Finnveden_1999, Pelletier_2014) can also render study findings unreliable or highly contingent.

+ P6L13ff: “An LCA study result is ultimately an assertion that, for some input data known as the “model,”…”  1) That describes Attributional LCA, please specify. 2) please rephrase the start of this sentence, as it is unclear how (input) data can be a model.

The comment has been updated to read: "An LCA study result is ultimately an assertion that for some constructed PSM, the delivery of a particular reference flow is associated with a certain amount of environmental impact or potential impact."  The applicability of the statement is not limited to attributional LCA.

+ P8L31: Please also cite DOI 10.1111/jiec.12142 (Majeau-Bettez et al. 2014) here for the link between supply and use tables and input-output models in LCA.

This is now done.

+ P8L40: “generated non-reference outputs (negative values)” Only for some constructs. In the case of partitioning allocation there are no negative inflows as input requirements are split so that only single output process descriptions result.

Noted in a footnote.

+ Equations 3 and 4 are not equivalent. Equ. 4 cannot be used for a structural path analysis, for example, as it contains aggregate information only. Please be more specific.

Updated to read "...equivalency of results derived from Eq.s 3 and 4..."

+ P11L22: “The foreground graph Af can be distilled into a vector of…” please use more scientific language here. A_f is not a graph but a description of a graph, and distilling reminds me of other things than science…

The passage has been removed.

+ Figure S1 provides a very good summary of the different aggregation steps; consider moving it to the main paper.

Thank you for the compliment and the suggestion; this has been done in the revision.

+ There are two A-matrices in LCA: One is the Suh and Heijungs matrix of flows and the other one is the matrix of technical coefficients. Please distinguish better between the two, e.g., on P12L45, where you write about flows in the A_f matrix, but it should read ‘coefficients in the A_f matrix’.

In the revision, only the Leontief matrix is used. Other passages have been updated or removed.

+ P13L31: Only aggregated results can be reproduced from the aggregated foreground. Please make this limitation explicit.

Noted: "The aggregated result can also be computed from the aggregated foreground..." (p. X l x)

+ P31L41: “Generally, any directed graph can be represented in Af ,…” That is not correct. Af can only contain bipartite graphs with a 1:1 correspondence between industries and nodes, that means, an industrial inventory (SUT) after allocation. Cf. DOI 10.1111/jiec.12306 for the argument.

Noted: "Generally, the foreground matrix $A_f$ can represent any process-flow model with a 1:1 correspondence between processes and flows, i.e. any allocated supply-and-use table..."

+ Figures 4 and 5: Units are missing. Is the rightmost column the row sum of the detailed exchange table? Please provide more explanations in the Table caption. The symbols A_f, B_f, etc. that are frequently used in the text should be displayed here as well.

These exhibits have been moved into the supplementary materials.
NEED MORE FINISH HERE

### 2017-11-15 aggregation/agg-basic-lca.tex
Pauliuk comments

We begin by providing a mathematical formulation of the LCA problem.  An LCA computation is performed over a collection of linked unit processes, each of which describes a single operation or reference product.  The task of procedure can be described in two steps: making an inventory of environmental interventions; and determining their environmental impacts.  If the processes are assumed to be linear and time-invariant, the problem can be written in terms of matrix algebra.  There are two common formulations: the ``technology matrix'' approach and the ``input-output'' or Leontief approach \citep{Heijungs2002, Pauliuk_2015_framework}.  Although the two models are understood to be equivalent, the input-output approach is computationally more efficient in some applications because it lends itself to iterative solution \citep{Peters_JLCA_2007}.  That formulation may be stated:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the \emph{direct requirements matrix}, where each column of values reports the necessary inputs (positive values) and generated non-reference outputs (negative values)%\footnote{Whether negative values are present or allowed depends on the particular construct or co-production method being used. \cite{Majeau_Bettez_2014}}
per unit of a process's reference output; $B$ is the environmental intervention or emission matrix; and $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions with respect to a given impact assessment method.  In this model, $\mathbf{y}$ is the externally specified final demand, which is the product or service whose delivery is to be assessed, and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.

>>>>

We begin by providing a mathematical formulation of the LCA problem.  An LCA computation is performed over a collection of linked unit processes, each of which describes a single operation or reference product.  The task of procedure can be described in two steps: making an inventory of environmental interventions; and determining their environmental impacts.  If the processes are assumed to be linear and time-invariant, the problem can be written in terms of matrix algebra.  There are two common formulations: the ``technology matrix'' approach and the ``input-output'' or Leontief approach \citep{Heijungs2002, Pauliuk_2015_framework}.  Although the two models are understood to be equivalent, the input-output approach is computationally more efficient in some applications because it lends itself to iterative solution \citep{Peters_JLCA_2007}.  That formulation may be stated:
\begin{equation}
s = \mathbf{e}^T \cdot B \cdot \left(I - A\right)^{-1} \cdot \mathbf{y}
\label{eqn:leontief}
\end{equation}
Here $A$ represents the \emph{direct requirements matrix}, where each column of values reports the necessary inputs (positive values) and generated non-reference outputs (negative values)\footnote{Whether negative values are present or allowed depends on the particular construct or co-production method being used. \cite{Majeau_Bettez_2014}}
per unit of a process's reference output; $B$ is the environmental intervention or emission matrix; and $\mathbf{e}$ is a column vector of characterization factors for the environmental emissions with respect to a given impact assessment method.  In this model, $\mathbf{y}$ is the externally specified final demand, which is the product or service whose delivery is to be assessed, and $s$ is the numerical impact score or category indicator.  For simplicity, we will consider the computation of a single impact result.  However, it is straightforward to imagine the more typical case, in which $E$ is a matrix of characterization vectors, and $\mathbf{s}$ is a vector of results.

### 2017-11-15 aggregation/agg-intro.tex
Pauliuk comments

Although there has been considerable effort to normalize and harmonize 
LCI database design methodology and data interchange \citep{UNEP_2011, JRC_ILCD_ELCD_2013, Mila_e_Canals_2015, Ingwersen_JLCA_2015}, less attention has been paid to harmonizing PSMs.  These models are often highly complex and include countless modeling decisions, approximations, and assumptions made by a study author \citep{Lloyd2007, reap2008_I}.  Consequently, while comparative evaluations can be made regarding alternative cases or scenarios within a single study, comparisons across studies are much more challenging \citep{Heath2012, Henriksson2014}.  Even when relatively simple and homogeneous systems are considered, wide variations can be found in the results of studies from different authors \citep{van_der_Harst_2013, Turconi_2013}.
A lack of transparency in reporting, particularly regarding definitions of study scope and system boundary,
is a crucial challenge to the interpretation of results in comparative analysis \citep{Cleary2009, Laurent_2014}.  A further challenge to transparency can be found in the use of input data that are confidential or proprietary \citep{Kuczenski_2017}.  One consequence of all these challenges is the high cost and uneven rigor of critical review, in which the complexities of LCA come face to face with the limitations of current practice \citep{Curran2014}.

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
Other factors such as allocation strategy \citep{Pelletier_2014} can also render study findings unreliable or highly contingent.  Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.  In the scope of Environmental Product Declarations (EPDs), studies within an industry are supposed to be rendered comparable by adhering to a common product category rule (PCR) \citep{Fet_2006}.  However, this prescription is insufficient to ensure the comparability of different declarations, even if they use the same PCR \citep{Modahl_2012}.  In addition, PCRs are often themselves not unique and can include wide variations in scope and boundary requirements \citep{Subramanian_2012}, leading to ambiguity.

>>>>

Although there has been considerable effort to normalize and harmonize 
LCI database design methodology and data interchange \citep{UNEP_2011, JRC_ILCD_ELCD_2013, Mila_e_Canals_2015, Ingwersen_JLCA_2015}, less attention has been paid to harmonizing PSMs.  These models are often highly complex and include countless modeling decisions, approximations, and assumptions made by a study author \citep{Lloyd2007, reap2008_I}.
Factors such as co-production strategy \citep{Finnveden_1999, Pelletier_2014} can also render study findings unreliable or highly contingent.  Consequently, while comparative evaluations can be made regarding alternative cases or scenarios within a single study, comparisons across studies are much more challenging \citep{Heath2012, Henriksson2014}.  Even when relatively simple and homogeneous systems are considered, wide variations can be found in the results of studies from different authors \citep{van_der_Harst_2013, Turconi_2013}.
A lack of transparency in reporting, particularly regarding definitions of study scope and system boundary,
is a crucial challenge to the interpretation of results in comparative analysis \citep{Cleary2009, Laurent_2014}.  A further challenge to transparency can be found in the use of input data that are confidential or proprietary \citep{Kuczenski_ESD_2017}.  One consequence of all these challenges is the high cost and uneven rigor of critical review, in which the complexities of LCA come face to face with the limitations of current practice \citep{Curran2014}.

As LCA gains prominence, particularly in the policy realm, these problems become more acute.  If LCA is to be used in policy, it will necessarily have the effect of recommending certain product systems, technologies, or approaches at the expense of others; yet in that event it is critically important that there be a consensus among stakeholders that the quantitative results are credible and well-supported \citep{Rainville_2015, McManus_2015}.  As noted above, this level of consensus is hard to realize, in part because of the conflicting implications inherent in different analytic modes.
Plevin et al.\ famously observed in the strongest terms that careless reporting of attributional LCA study results can distort the significance of findings in a policy context \citep{Plevin_2013}, a shortcoming that can be found in other modes as well \citep{Brandao_2014}.  In the scope of Environmental Product Declarations (EPDs), studies within an industry are supposed to be rendered comparable by adhering to a common product category rule (PCR) \citep{Fet_2006}.  However, this prescription is insufficient to ensure the comparability of different declarations, even if they use the same PCR \citep{Modahl_2012}.  In addition, PCRs are often themselves not unique and can include wide variations in scope and boundary requirements \citep{Subramanian_2012}, leading to ambiguity.
